\vspace{-0.2in}
Current interest in advanced nuclear energy and \gls{MSR} concepts has enhanced demand in building the tools to analyze these systems. This thesis introduces a Python script, SaltProc, which simulates \gls{MSR} online reprocessing by modeling the changing isotopic composition of an irradiated fuel salt. SaltProc couples with the Monte Carlo code, SERPENT 2, for neutron transport and depletion calculations. SaltProc capabilities include a generic geometry capable of modeling multi-region and multi-flow systems, time-dependent feeds and removals, and specific separation efficiency for each element or isotope removal flow. Generally applicable  capabilities are illuminated in this thesis in three applied problems: (1) simulating the startup of a thorium-fueled \gls{MSBR} fuel cycle to find equilibrium fuel composition (when multiplication factor of the full-core model and the $^{233}$U concentration in the fuel salt are invariant in time); (2) determining the effect of the fuel salt irradiation with online reprocessing on \gls{MSBR} operations; and (3) estimating \gls{MSBR} fuel cycle performance by computing the $^{232}$Th feed rate over 20 years of operation and comparing with available data.

In the first application, full-core depletion in the \gls{MSBR} demonstrated that (1) equilibrium fuel composition could be achieved after 16 years and (2) the multiplication factor stabilizes after 6 years of operation. In the second application, fuel salt irradiation with simulated fission product removal and fissile/fertile feed causes considerable neutron energy spectrum hardening; this spectral shift has a problematic impact on safety parameters (e.g., temperature reactivity feedback, reactivity control system worth). In the third application, the average $^{232}$Th feed rate throughout 20 years of operation is 2.39 kg/day or 100 g/GWh$_e$ which is a good agreement with other recent research. Problematic effects of neutron energy spectrum hardening during \gls{MSBR} operation should be taken into account for neutronics, multi-physics, and fuel cycle performance analysis.